---
title: "Tarea 5"
author: "Sonya Leilany Lorea Aguilera"
date: '2025'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(discrtr)
library(dplyr)
library(ggplot2)
library(mlogit)
library(readr)
library(gplots)
library(ggmosaic)
library(treemapify)
library(ggridges)
library(ggalluvial)
library(evd)
library(htmlwidgets)
library(kableExtra)
library(plotly)
library(tidyr)
library(evd)
```

```{r}
data("Heating")
glimpse(Heating)
model3 <- mlogit(depvar ~ ic + oc, Heating, shape = "wide", choice = "depvar", reflevel = "ec", varying = c(3:12))
ic_min <- Heating[1, ]
ic_mean <- Heating[1, ]
ic_max <- Heating[1, ]
min_cost <- Heating %>% select(starts_with("ic")) %>%
                        summarise(across(.cols = everything(), min))

mean_cost <- Heating %>% select(starts_with("ic")) %>%
             summarise(across(.cols = everything(), mean))

max_cost <- Heating %>% select(starts_with("ic")) %>%  
                        summarise(across(.cols = everything(), max))
ic_mean[3:12] <- mean_cost
```


1. What is the difference between a marginal effect and an elasticity?
La elasticidad mide los cambios de una variable dependiente con respecto en los cambios porcentuales de otra variable, mientras que el efecto marginal mide los cambios de una variable con respecto a los cambios en una unidad de la otra variable. Es decir la elasticidad mide el cambio porcentual y el cambio marginal es en unidades.
 
2. Why is it inappropriate to calculate the elasticity of a dummy variable?
Las variables dummy son una forma de categorizar variables que generalmente solo tienen dos alternativa de respuesta, como sí/no, género, etc, se les suele asignar un valor de 1 y 0, por tanto no tienen una escala continua y no tendría sentido calcular cambios porcentuales entre variables.
 
3. Use Model 3 in this chapter and calculate the marginal effects and the elasticities for operating cost at the mean of all variables.

## Efectos marginales
```{r}
effects(model3, covariate = "ic", type = "ra", data = mlogit.data(ic_mean, shape = "wide",
        choice = "depvar", varying = 3:12))
```
## Elasticidad
```{r}
effects(model3, covariate = "ic", type = "rr", data = mlogit.data(ic_mean, shape = "wide",
        choice = "depvar", varying = 3:12))
```

4. Use Model 3 in this chapter to calculate the rebate needed to reach a 10% penetration rate of heat pumps. Estimate a new model that extends Model 3 by introducing the age of the household head. Use the electric room system (“er”) as the reference level.
 
```{r}
H_rebate <- Heating %>% mutate(ic.er=0.947*ic.er)

data.frame(Politica = c("Sin reembolso", "5.3% de descuento"),
           rbind(apply(predict(model3, newdata = mlogit.data(Heating, shape = "wide", choice = "depvar",
                       varying = c(3:12))), 2, mean),
                 apply(predict(model3, newdata = mlogit.data(H_rebate, shape = "wide", choice = "depvar",
                       varying = c(3:12))), 2, mean)))
```
Se requiere cerca de un 5.3% de descuento para lograr un 10% de penetración.

```{r}
H_rebate_agehed <- Heating %>% mutate(ic.er=0.946*ic.er)


model4 <- mlogit(depvar ~ ic + oc | agehed, Heating, shape = "wide", choice = "depvar", reflevel = "ec", varying = c(3:12))


data.frame(Politica = c("Sin reembolso", "5.4% de descuento"),
           rbind(apply(predict(model4, newdata = mlogit.data(Heating, shape = "wide", choice = "depvar",
                       varying = c(3:12))), 2, mean),
                 apply(predict(model4, newdata = mlogit.data(H_rebate_agehed, shape = "wide", choice = "depvar",
                       varying = c(3:12))), 2, mean)))
```
Se requiere cerca de un 5.4% de descuento para lograr un 10% de penetración.
 
5. Use the likelihood ratio test to compare your new model to Model 3. Discuss the results.
 
```{r}
lrtest(model3, model4)
```

6. Is the ratio of the coefficient of installation (or operation) cost to the coefficient of age of household head meaningful? Explain.

Con base en los resultados del likelihood ratio test, es significativo en 0.1, es decir, es poco significativo.

