---
title: "Tarea 1; Data, Models and Softwares"
author: "Sonya Leilany Lorea Aguilera"
date: "2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(discrtr) 
library(mlogit) 
library(readr) 
library(plotly)
```


###Invoke data set Mode from package mlogit. To do this you need to first load the package. This is a data set with choices about mode of transportation. This is done as follows:

```{r}
data("Mode")
```

###Once you have loaded the data set, answer the following questions:
Describe this data set. How many variables are there and of which type (i.e., categorical/quantitative)?

```{r}
glimpse(Mode)
```
Variables categoricas: 1="Choice"
Variables cuantitativas: 8=Todas las demás


###How many different modes of transportation are in this data set? What is the most popular mode? What is the least popular mode?

```{r}
summary(Mode)
```
```{r}
ggplot(Mode)+
  geom_bar(aes(choice, fill=choice), color="black", show.legend = FALSE)+
  theme_bw()
```
Con base en la variable "choice", existen 4 formas de transporte, la más popular es el carro y la menos popular es el carpool.

###In general, what is the most expensive mode? The least expensive?

```{r}
ggplot(Mode)+
  geom_density(aes(cost.car, fill="car"), color="black", alpha=0.5)+
  geom_density(aes(cost.carpool , fill="carpool"), color="black", alpha=0.5)+
  geom_density(aes(cost.bus, fill="bus"), color="black", alpha=0.5)+
  geom_density(aes(cost.rail, fill="rail"), color="black", alpha=0.5)+
  scale_fill_manual(values=c("car"= "firebrick", "carpool"="dodgerblue", "bus"="darkgoldenrod2", "rail"="cyan"), name="Mode")+
  xlab("Cost")+
theme_bw()
  
```
La opción más costosa es el carro, y la menos costosa es el camión.

###Create a plot showing the univariate distributions of time by car and time by bus. Discuss.
```{r}
grafica_time <- ggplot(Mode)+
  geom_density(aes(time.car, fill="car"), color="black", alpha=0.5)+
  #geom_density(aes(time.carpool , fill="carpool"), color="black", alpha=0.5)+
  geom_density(aes(time.bus, fill="bus"), color="black", alpha=0.5)+
  #geom_density(aes(time.rail, fill="rail"), color="black", alpha=0.5)+
  #scale_fill_manual(values=c("car"= "firebrick", "carpool"="dodgerblue", "bus"="darkgoldenrod2", "rail"="cyan"), name="Mode")+
  scale_fill_manual(values=c("car"= "firebrick", "bus"="darkgoldenrod2"), name="Mode")+
  xlab("Time")+
theme_bw()

ggplotly(grafica_time)  
```
El tiempo estimado para un carro es superior al tiempo de transportarse en autobus.

###How do choices relate to cost by the different modes?
```{r}
varnum <- Mode |> dplyr::select(where(is.numeric))
varnum <- names(varnum)

for (var in varnum) {
grafica_box <- ggplot(Mode)+
  geom_boxplot(aes(choice, eval(as.name(var)), fill=choice), show.legend=FALSE)+
  ylab(var)+
  theme_bw()
  
  print(grafica_box)
    
}

```
Según que tan costoso sea un transporte u otro, hay mayor preferencia por un modo u otro, es decir, mientras más caro sea el modo de transporte, menor cantidad de población lo escoge.

