---
title: "Tarea 2"
author: "Sonya Leilany Lorea Aguilera"
date: '2025'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(discrtr)
library(mlogit)
library(readr) 
library(gplots) 
library(ggmosaic) 
library(treemapify)
library(ggridges)
library(ggalluvial)
library(plotly)
```

Se carga el conjunto de datos, se muestran las variables y su tipo:

```{r}
data(Car)
glimpse(Car)
```

En este caso es necesario añadir columnas acerca de la elección, las gráficas se realizarán con la data frame `new_car`:

```{r}
new_car <- Car
choice_char <- gsub("choice","", new_car$choice)
new_car <- new_car |> mutate(choice_char = choice_char)

type <- vector("character", nrow(new_car))
fuel <- vector("character", nrow(new_car))
price <- vector("numeric", nrow(new_car))
range <- vector("numeric", nrow(new_car))
acc <- vector("numeric", nrow(new_car))
speed <- vector("numeric", nrow(new_car))
pollution <- vector("numeric", nrow(new_car))
size <- vector("numeric", nrow(new_car))
space <- vector("numeric", nrow(new_car))
cost <- vector("numeric", nrow(new_car))
station <- vector("numeric", nrow(new_car))
for (i in 1:nrow(new_car)) {
  type[i] <- as.character(new_car[i, paste0("type", new_car$choice_char[i], collapse = "")])
  fuel[i] <- as.character(new_car[i, paste0("fuel", new_car$choice_char[i], collapse = "")])
  price[i] <- new_car[i, paste0("price", new_car$choice_char[i], collapse = "")]
  range[i] <- new_car[i, paste0("range", new_car$choice_char[i], collapse = "")]
  acc[i] <- new_car[i, paste0("acc", new_car$choice_char[i], collapse = "")]
  speed[i] <- new_car[i, paste0("speed", new_car$choice_char[i], collapse = "")]
  pollution[i] <- new_car[i, paste0("pollution", new_car$choice_char[i], collapse = "")]
  size[i] <- new_car[i, paste0("size", new_car$choice_char[i], collapse = "")]
  space[i] <- new_car[i, paste0("space", new_car$choice_char[i], collapse = "")]
  cost[i] <- new_car[i, paste0("cost", new_car$choice_char[i], collapse = "")]
  station[i] <- new_car[i, paste0("station", new_car$choice_char[i], collapse = "")]
}

new_car <- new_car |> mutate(choice_num= factor(choice_char), type=type, fuel=fuel, price=price, range=range,
                             acc=acc, speed=speed, pollution=pollution, size=size, space=space, cost=cost,
                             station=station)

new_car$type <- factor(new_car$type, levels = levels(new_car$type1))
new_car$fuel <- factor(new_car$fuel, levels = levels(new_car$fuel1))
```

* How many variables are there in this data set and of which type (i.e., categorical/quantitative)?
Variables categoricas: Todas las demás
Variables cuantitativas: precio, velocidad y pollution

* Choose four relevant categorical variables from this data set and describe them using univariate, bivariate, and multivariate techniques. Discuss your results.

```{r}
new_car |> ggplot(aes(x = fuel)) + 
  geom_bar(color = "black", fill = "white") + 
  labs(x = "Tipo de vehiculos", y = "Number of respondents")
```

```{r}
new_car |> group_by(fuel,type) |> 
  summarize(n = n(), .groups = "drop") |> 
  ggplot(aes(x = fuel, y = type)) +
  geom_tile(aes(fill = n)) +
  labs(x = "fuel", y = "type", fill = "Number of respondents")
```


```{r}
new_car |>  ggplot(aes(x = type, y = speed, fill = type)) +
  geom_violin(trim = TRUE) +
  geom_boxplot(width = 0.1, fill = "white") + 
  labs(x="type", y = expression("speed"), fill = "Mode")
```


* What ideas about individuals’ choices regarding car do you develop from this EDA exercise?
Hay una mayor cantidad de personas que tienen vehiculos electricos.

